---
title: "Untitled"
date: "29/01/2020"
output: html_document
---

find the yield curve
```{r setup, include=FALSE}
library(readxl)
bonds_final <- read_excel("Desktop/bonds_final.xlsx")
attach(bonds_final)
#find dirty price#
date02=as.Date(c("2020-01-02","2019-09-01"))
date03=as.Date(c("2020-01-03","2019-09-01"))
date06=as.Date(c("2020-01-06","2019-09-01"))
date07=as.Date(c("2020-01-07","2019-09-01"))
date08=as.Date(c("2020-01-08","2019-09-01"))
date09=as.Date(c("2020-01-09","2019-09-01"))
date10=as.Date(c("2020-01-10","2019-09-01"))
date13=as.Date(c("2020-01-13","2019-09-01"))
date14=as.Date(c("2020-01-14","2019-09-01"))
date15=as.Date(c("2020-01-15","2019-09-01"))

date2=as.Date(c("2020-01-02","2019-12-01"))
date3=as.Date(c("2020-01-03","2019-12-01"))
date6=as.Date(c("2020-01-06","2019-12-01"))
date7=as.Date(c("2020-01-07","2019-12-01"))
date8=as.Date(c("2020-01-08","2019-12-01"))
date9=as.Date(c("2020-01-09","2019-12-01"))
date10a=as.Date(c("2020-01-10","2019-12-01"))
date13a=as.Date(c("2020-01-13","2019-12-01"))
date14a=as.Date(c("2020-01-14","2019-12-01"))
date15a=as.Date(c("2020-01-15","2019-12-01"))

diff02=c(as.numeric(date02[1]-date02[2]),as.numeric(date02[1]-date02[2]),as.numeric(date02[1]-date02[2]),as.numeric(date02[1]-date02[2]),as.numeric(date02[1]-date02[2]),as.numeric(date2[1]-date2[2]),as.numeric(date02[1]-date02[2]),as.numeric(date2[1]-date2[2]),as.numeric(date02[1]-date02[2]),as.numeric(date02[1]-date02[2]),as.numeric(date02[1]-date02[2]))
diff03=c(as.numeric(date03[1]-date03[2]),as.numeric(date03[1]-date03[2]),as.numeric(date03[1]-date03[2]),as.numeric(date03[1]-date03[2]),as.numeric(date03[1]-date03[2]),as.numeric(date3[1]-date3[2]),as.numeric(date03[1]-date03[2]),as.numeric(date3[1]-date3[2]),as.numeric(date03[1]-date03[2]),as.numeric(date03[1]-date03[2]),as.numeric(date03[1]-date03[2]))
diff06=c(as.numeric(date06[1]-date06[2]),as.numeric(date06[1]-date06[2]),as.numeric(date06[1]-date06[2]),as.numeric(date06[1]-date06[2]),as.numeric(date06[1]-date06[2]),as.numeric(date6[1]-date6[2]),as.numeric(date06[1]-date06[2]),as.numeric(date6[1]-date6[2]),as.numeric(date06[1]-date06[2]),as.numeric(date06[1]-date06[2]),as.numeric(date06[1]-date06[2]))
diff07=c(as.numeric(date07[1]-date07[2]),as.numeric(date07[1]-date07[2]),as.numeric(date07[1]-date07[2]),as.numeric(date07[1]-date07[2]),as.numeric(date07[1]-date07[2]),as.numeric(date7[1]-date7[2]),as.numeric(date07[1]-date07[2]),as.numeric(date7[1]-date7[2]),as.numeric(date07[1]-date07[2]),as.numeric(date07[1]-date07[2]),as.numeric(date07[1]-date07[2]))
diff08=c(as.numeric(date08[1]-date08[2]),as.numeric(date08[1]-date08[2]),as.numeric(date08[1]-date08[2]),as.numeric(date08[1]-date08[2]),as.numeric(date08[1]-date08[2]),as.numeric(date8[1]-date8[2]),as.numeric(date08[1]-date08[2]),as.numeric(date8[1]-date8[2]),as.numeric(date08[1]-date08[2]),as.numeric(date08[1]-date08[2]),as.numeric(date08[1]-date08[2]))
diff09=c(as.numeric(date09[1]-date09[2]),as.numeric(date09[1]-date09[2]),as.numeric(date09[1]-date09[2]),as.numeric(date09[1]-date09[2]),as.numeric(date09[1]-date09[2]),as.numeric(date9[1]-date9[2]),as.numeric(date09[1]-date09[2]),as.numeric(date9[1]-date9[2]),as.numeric(date09[1]-date09[2]),as.numeric(date09[1]-date09[2]),as.numeric(date09[1]-date09[2]))
diff10=c(as.numeric(date10[1]-date10[2]),as.numeric(date10[1]-date10[2]),as.numeric(date10[1]-date10[2]),as.numeric(date10[1]-date10[2]),as.numeric(date10[1]-date10[2]),as.numeric(date10a[1]-date10a[2]),as.numeric(date10[1]-date10[2]),as.numeric(date10a[1]-date10a[2]),as.numeric(date10[1]-date10[2]),as.numeric(date10[1]-date10[2]),as.numeric(date10[1]-date10[2]))
diff13=c(as.numeric(date13[1]-date13[2]),as.numeric(date13[1]-date13[2]),as.numeric(date13[1]-date13[2]),as.numeric(date13[1]-date13[2]),as.numeric(date13[1]-date13[2]),as.numeric(date13a[1]-date13a[2]),as.numeric(date13[1]-date13[2]),as.numeric(date13a[1]-date13a[2]),as.numeric(date13[1]-date13[2]),as.numeric(date13[1]-date13[2]),as.numeric(date13[1]-date13[2]))
diff14=c(as.numeric(date14[1]-date14[2]),as.numeric(date14[1]-date14[2]),as.numeric(date14[1]-date14[2]),as.numeric(date14[1]-date14[2]),as.numeric(date14[1]-date14[2]),as.numeric(date14a[1]-date14a[2]),as.numeric(date14[1]-date14[2]),as.numeric(date14a[1]-date14a[2]),as.numeric(date14[1]-date14[2]),as.numeric(date14[1]-date14[2]),as.numeric(date14[1]-date14[2]))
diff15=c(as.numeric(date15[1]-date15[2]),as.numeric(date15[1]-date15[2]),as.numeric(date15[1]-date15[2]),as.numeric(date15[1]-date15[2]),as.numeric(date15[1]-date15[2]),as.numeric(date15a[1]-date15a[2]),as.numeric(date15[1]-date15[2]),as.numeric(date15a[1]-date15a[2]),as.numeric(date15[1]-date15[2]),as.numeric(date15[1]-date15[2]),as.numeric(date15[1]-date15[2]))

dir02=Jan_02+(diff02/365)*(`Coupon_in_%`)
dir03=Jan_03+(diff03/365)*(`Coupon_in_%`)
dir06=Jan_06+(diff06/365)*(`Coupon_in_%`)
dir07=Jan_07+(diff07/365)*(`Coupon_in_%`)
dir08=Jan_08+(diff08/365)*(`Coupon_in_%`)
dir09=Jan_09+(diff09/365)*(`Coupon_in_%`)
dir10=Jan_10+(diff10/365)*(`Coupon_in_%`)
dir13=Jan_13+(diff13/365)*(`Coupon_in_%`)
dir14=Jan_14+(diff14/365)*(`Coupon_in_%`)
dir15=Jan_15+(diff15/365)*(`Coupon_in_%`)
bonds_dir=cbind(bonds_final,dir02,dir03,dir06,dir07,dir08,dir09,dir10,dir13,dir14,dir15)
dirpricelist=data.frame(dir02,dir03,dir06,dir07,dir08,dir09,dir10,dir13,dir14,dir15)

library(jrvFinance)
#find yields to maturity
x=rep(1)
#for each day yields to maturity of bonds
Ytm02=data.frame(x,x,x,x,x,x,x,x,x,x,x)
Ytm03=data.frame(x,x,x,x,x,x,x,x,x,x,x)
Ytm06=data.frame(x,x,x,x,x,x,x,x,x,x,x)
Ytm07=data.frame(x,x,x,x,x,x,x,x,x,x,x)
Ytm08=data.frame(x,x,x,x,x,x,x,x,x,x,x)
Ytm09=data.frame(x,x,x,x,x,x,x,x,x,x,x)
Ytm10=data.frame(x,x,x,x,x,x,x,x,x,x,x)
Ytm13=data.frame(x,x,x,x,x,x,x,x,x,x,x)
Ytm14=data.frame(x,x,x,x,x,x,x,x,x,x,x)
Ytm15=data.frame(x,x,x,x,x,x,x,x,x,x,x)

#calculate yield to maturity for each day
coupon=`Coupon_in_%`/100
mature=as.Date(Maturity_date)

for (i in c(1:11))
{Ytm02[i]=bond.yield("2020-01-02",mature[i],coupon[i],freq=2,price=Jan_02[i])}
Ytm02=as.numeric(Ytm02)

for (i in c(1:11))
{Ytm03[i]=bond.yield("2020-01-03",mature[i],coupon[i],freq=2,price=Jan_03[i])}
Ytm03=as.numeric(Ytm03)

for (i in c(1:11))
{Ytm06[i]=bond.yield("2020-01-06",mature[i],coupon[i],freq=2,price=Jan_06[i])}
Ytm06=as.numeric(Ytm06)

for (i in c(1:11))
{Ytm07[i]=bond.yield("2020-01-07",mature[i],coupon[i],freq=2,price=Jan_07[i])}
Ytm07=as.numeric(Ytm07)

for (i in c(1:11))
{Ytm08[i]=bond.yield("2020-01-08",mature[i],coupon[i],freq=2,price=Jan_08[i])}
Ytm08=as.numeric(Ytm08)

for (i in c(1:11))
{Ytm09[i]=bond.yield("2020-01-09",mature[i],coupon[i],freq=2,price=Jan_09[i])}
Ytm09=as.numeric(Ytm09)

for (i in c(1:11))
{Ytm10[i]=bond.yield("2020-01-10",mature[i],coupon[i],freq=2,price=Jan_10[i])}
Ytm10=as.numeric(Ytm10)

for (i in c(1:11))
{Ytm13[i]=bond.yield("2020-01-13",mature[i],coupon[i],freq=2,price=Jan_13[i])}
Ytm13=as.numeric(Ytm13)

for (i in c(1:11))
{Ytm14[i]=bond.yield("2020-01-14",mature[i],coupon[i],freq=2,price=Jan_14[i])}
Ytm14=as.numeric(Ytm14)

for (i in c(1:11))
{Ytm15[i]=bond.yield("2020-01-15",mature[i],coupon[i],freq=2,price=Jan_15[i])}
Ytm15=as.numeric(Ytm15)

#time to maturity for each day
dates <- c("2020-01-02", "2020-01-03", "2020-01-06", "2020-01-07", "2020-01-08", "2020-01-09", "2020-01-10", "2020-01-13", "2020-01-14", "2020-01-15")
ttmJan02=yearFraction(dates[1],mature)
ttmJan03=yearFraction(dates[2],mature)
ttmJan06=yearFraction(dates[3],mature)
ttmJan07=yearFraction(dates[4],mature)
ttmJan08=yearFraction(dates[5],mature)
ttmJan09=yearFraction(dates[6],mature)
ttmJan10=yearFraction(dates[7],mature)
ttmJan13=yearFraction(dates[8],mature)
ttmJan14=yearFraction(dates[9],mature)
ttmJan15=yearFraction(dates[10],mature)

#plot the yield to maturity
plot(ttmJan02, Ytm02, type = "b", frame = FALSE, pch = 19, col="red", xlab = "time to maturity in years", ylab="yield")
lines(ttmJan03, Ytm03, pch = 18, col = "blue", type = "b", lty = 1)
lines(ttmJan06, Ytm06, pch = 18, col = "yellow", type = "b", lty = 1)
lines(ttmJan07, Ytm07, pch = 18, col = "green", type = "b", lty = 1)
lines(ttmJan08, Ytm08, pch = 18, col = "brown", type = "b", lty = 1)
lines(ttmJan09, Ytm09, pch = 18, col = "purple", type = "b", lty = 1)
lines(ttmJan10, Ytm10, pch = 18, col = "black", type = "b", lty = 1)
lines(ttmJan13, Ytm13, pch = 18, col = "grey", type = "b", lty = 1)
lines(ttmJan14, Ytm14, pch = 18, col = "pink", type = "b", lty = 1)
lines(ttmJan15, Ytm15, pch = 18, col = "orange", type = "b", lty = 1)

legend("topright", legend=c("2020-01-02", "2020-01-03","2020-01-06","2020-01-07","2020-01-08","2020-01-09","2020-01-10","2020-01-13","2020-01-14","2020-01-15"),
       col=c("red", "blue","yellow","green","brown","purple","black","grey","pink","orange"), lty = 1, cex=0.5)


```






find the spot rate 
```{r setup, include=FALSE}
#find spot rate
payoff_times = c(1,2, 3, 4, 5, 5, 7, 7, 9, 10, 11)
dates <- c("2020-01-02", "2020-01-03", "2020-01-06", "2020-01-07", "2020-01-08", "2020-01-09", "2020-01-10", "2020-01-13", "2020-01-14", "2020-01-15")
time_since_last_payment=c(123,124,127,128,129,130,131,134,135,136)

spotrate=data.frame(x,x,x,x,x,x,x,x,x,x)

# calculate spot rate for bond1 in each day(mature in '2020-03-01')
cashflow_bond1=c()
for(i in c(1:10))
  cashflow_bond1=append(cashflow_bond1, bond.TCF(dates[i],mature[1],coupon[1],freq = 2,redemption_value =100)$cf)
spotratebond1=c()
for (i in c(1:10)) 
  spotratebond1 = append(spotratebond1,log(cashflow_bond1[i]/dirpricelist[[i]][1])/yearFraction(dates[i],mature[1]))


## calculate spot rate for bond2 in each day(mature in '2020-09-01')
cashflow_bond2=bond.TCF(dates[i],mature[2],coupon[2],freq = 2,redemption_value =100)$cf

spotratebond2=c()
for (i in c(1:10)) 
  spotratebond2 = append(spotratebond2,log(cashflow_bond2[2]/(dirpricelist[[i]][2]- cashflow_bond2[1] * exp(-1*spotratebond1[i] * yearFraction(dates[i],mature[1]))))/yearFraction(dates[i],mature[2]))

## calculate spot rate for bond3 in each day(mature in '2021-03-01')
cashflow_bond3=bond.TCF(dates[i],mature[3],coupon[3],freq = 2,redemption_value =100)$cf
spotratebond3=c()
for (i in c(1:10)) 
  spotratebond3 = append(spotratebond3,log(cashflow_bond3[3]/(dirpricelist[[i]][3]- cashflow_bond3[1] * exp(-1*spotratebond1[i] * yearFraction(dates[i],mature[1]))-cashflow_bond3[2]*exp(-1*spotratebond2[i]*yearFraction(dates[i],mature[2]))))/yearFraction(dates[i],mature[3]))

## calculate spot rate for bond4 in each day(mature in '2021-09-01')
cashflow_bond4=bond.TCF(dates[i],mature[4],coupon[4],freq = 2,redemption_value =100)$cf
spotratebond4=c()
for (i in c(1:10)) 
  spotratebond4 = append(spotratebond4,log(cashflow_bond4[4]/(dirpricelist[[i]][4]- cashflow_bond4[1] * exp(-1*spotratebond1[i] * yearFraction(dates[i],mature[1]))-cashflow_bond4[2]*exp(-1*spotratebond2[i]*yearFraction(dates[i],mature[2]))-cashflow_bond4[3]*exp(-1*spotratebond3[i]*yearFraction(dates[i],mature[3]))))/yearFraction(dates[i],mature[4]))

## calculate spot rate for bond5 in each day(mature in '2022-03-01')
cashflow_bond5=bond.TCF(dates[i],mature[5],coupon[5],freq = 2,redemption_value =100)$cf
spotratebond5=c()
for (i in c(1:10)) 
  spotratebond5 = append(spotratebond5,log(cashflow_bond5[5]/(dirpricelist[[i]][5]- cashflow_bond5[1] * exp(-1*spotratebond1[i] * yearFraction(dates[i],mature[1]))-cashflow_bond5[2]*exp(-1*spotratebond2[i]*yearFraction(dates[i],mature[2]))-cashflow_bond5[3]*exp(-1*spotratebond3[i]*yearFraction(dates[i],mature[3]))-cashflow_bond5[4]*exp(-1*spotratebond4[i]*yearFraction(dates[i],mature[4]))))/yearFraction(dates[i],mature[5]))

## We need to do bond7 first(mature in '2023-03-01')
cashflow_bond7=bond.TCF(dates[i],mature[7],coupon[7],freq = 2,redemption_value =100)$cf
spotratebond7=c()
for (i in c(1:10)) 
  spotratebond7 = append(spotratebond7,log(cashflow_bond7[7]/(dirpricelist[[i]][7]- cashflow_bond7[1] * exp(-1*spotratebond1[i] * yearFraction(dates[i],mature[1]))-cashflow_bond7[2]*exp(-1*spotratebond2[i]*yearFraction(dates[i],mature[2]))-cashflow_bond7[3]*exp(-1*spotratebond3[i]*yearFraction(dates[i],mature[3]))-cashflow_bond7[4]*exp(-1*spotratebond4[i]*yearFraction(dates[i],mature[4]))-cashflow_bond7[5]*exp(-1*spotratebond5[i]*yearFraction(dates[i],mature[5]))-cashflow_bond7[6]*exp(-1*spotratebond5[i]*yearFraction(dates[i],'2022-09-01'))))/yearFraction(dates[i],mature[7]))

## bond6 by interpolation(mature in '2022-09-01')
spotratebond6=c()
for (i in c(1:10)) 
  spotratebond6 = append(spotratebond6,spotratebond5[i]+(spotratebond7[i]-spotratebond5[i])*yearFraction(mature[5],'2022-09-01')/yearFraction(mature[5],mature[7]))

## calculate spot rate for bond9 in each day(mature in '2024-03-01')
cashflow_bond9=bond.TCF(dates[i],mature[9],coupon[9],freq = 2,redemption_value =100)$cf
spotratebond9=c()
for (i in c(1:10)) 
  spotratebond9 = append(spotratebond9,log(cashflow_bond9[9]/(dirpricelist[[i]][9]- cashflow_bond9[1] * exp(-1*spotratebond1[i] * yearFraction(dates[i],mature[1]))-cashflow_bond9[2]*exp(-1*spotratebond2[i]*yearFraction(dates[i],mature[2]))-cashflow_bond9[3]*exp(-1*spotratebond3[i]*yearFraction(dates[i],mature[3]))-cashflow_bond9[4]*exp(-1*spotratebond4[i]*yearFraction(dates[i],mature[4]))-cashflow_bond9[5]*exp(-1*spotratebond5[i]*yearFraction(dates[i],mature[5]))-cashflow_bond9[6]*exp(-1*spotratebond6[i]*yearFraction(dates[i],'2022-09-01'))-cashflow_bond9[7]*exp(-1*spotratebond7[i]*yearFraction(dates[i],mature[7]))-cashflow_bond9[8]*exp(-1*spotratebond7[i]*yearFraction(dates[i],'2023-09-01'))))/yearFraction(dates[i],mature[9]))

## calculate spot rate for bond8 in each day(mature in '2023-09-01') by linear interpolation.
spotratebond8=c()
for (i in c(1:10)) 
  spotratebond8 = append(spotratebond8,spotratebond7[i]+(spotratebond9[i]-spotratebond7[i])*yearFraction(mature[7],'2023-09-01')/yearFraction(mature[7],mature[9]))

## calculate spot rate for bond10 in each day(mature in '2024-09-01')
cashflow_bond10=bond.TCF(dates[i],mature[10],coupon[10],freq = 2,redemption_value =100)$cf
spotratebond10=c()
for (i in c(1:10)) 
  spotratebond10 = append(spotratebond10,log(cashflow_bond10[10]/(dirpricelist[[i]][10]- cashflow_bond10[1] * exp(-1*spotratebond1[i] * yearFraction(dates[i],mature[1]))-cashflow_bond10[2]*exp(-1*spotratebond2[i]*yearFraction(dates[i],mature[2]))-cashflow_bond10[3]*exp(-1*spotratebond3[i]*yearFraction(dates[i],mature[3]))-cashflow_bond10[4]*exp(-1*spotratebond4[i]*yearFraction(dates[i],mature[4]))-cashflow_bond10[5]*exp(-1*spotratebond5[i]*yearFraction(dates[i],mature[5]))-cashflow_bond10[6]*exp(-1*spotratebond6[i]*yearFraction(dates[i],'2022-09-01'))-cashflow_bond10[7]*exp(-1*spotratebond7[i]*yearFraction(dates[i],mature[7]))-cashflow_bond10[8]*exp(-1*spotratebond8[i]*yearFraction(dates[i],'2023-09-01'))-cashflow_bond10[9]*exp(-1*spotratebond9[i]*yearFraction(dates[i],mature[9]))))/yearFraction(dates[i],mature[10]))

## calculate spot rate for bond11 in each day(mature in '2025-03-01')
cashflow_bond11=bond.TCF(dates[i],mature[11],coupon[11],freq = 2,redemption_value =100)$cf
spotratebond11=c()
for (i in c(1:10)) 
  spotratebond11 = append(spotratebond11,log(cashflow_bond11[11]/(dirpricelist[[i]][11]- cashflow_bond11[1] * exp(-1*spotratebond1[i] * yearFraction(dates[i],mature[1]))-cashflow_bond11[2]*exp(-1*spotratebond2[i]*yearFraction(dates[i],mature[2]))-cashflow_bond11[3]*exp(-1*spotratebond3[i]*yearFraction(dates[i],mature[3]))-cashflow_bond11[4]*exp(-1*spotratebond4[i]*yearFraction(dates[i],mature[4]))-cashflow_bond11[5]*exp(-1*spotratebond5[i]*yearFraction(dates[i],mature[5]))-cashflow_bond11[6]*exp(-1*spotratebond6[i]*yearFraction(dates[i],'2022-09-01'))-cashflow_bond11[7]*exp(-1*spotratebond7[i]*yearFraction(dates[i],mature[7]))-cashflow_bond11[8]*exp(-1*spotratebond8[i]*yearFraction(dates[i],'2023-09-01'))-cashflow_bond11[9]*exp(-1*spotratebond9[i]*yearFraction(dates[i],mature[9]))-cashflow_bond11[10]*exp(-1*spotratebond10[i]*yearFraction(dates[i],mature[10]))))/yearFraction(dates[i],mature[11]))

###spot rate for each day
spotratebond=list(spotratebond1,spotratebond2,spotratebond3,spotratebond4,spotratebond5,spotratebond6,spotratebond7,spotratebond8,spotratebond9,spotratebond10,spotratebond11)

###spot rate for Jan02
spotratejan02=c()
for(i in c(1:11))
  spotratejan02=append(spotratejan02,spotratebond[[i]][1])

###spot rate for Jan03
spotratejan03=c()
for(i in c(1:11))
  spotratejan03=append(spotratejan03,spotratebond[[i]][2])

###spot rate for Jan06
spotratejan06=c()
for(i in c(1:11))
  spotratejan06=append(spotratejan06,spotratebond[[i]][3])

###spot rate for Jan07
spotratejan07=c()
for(i in c(1:11))
  spotratejan07=append(spotratejan07,spotratebond[[i]][4])

###spot rate for Jan08
spotratejan08=c()
for(i in c(1:11))
  spotratejan08=append(spotratejan08,spotratebond[[i]][5])

###spot rate for Jan09
spotratejan09=c()
for(i in c(1:11))
  spotratejan09=append(spotratejan09,spotratebond[[i]][6])

###spot rate for Jan10
spotratejan10=c()
for(i in c(1:11))
  spotratejan10=append(spotratejan10,spotratebond[[i]][7])

###spot rate for Jan13
spotratejan13=c()
for(i in c(1:11))
  spotratejan13=append(spotratejan13,spotratebond[[i]][8])

###spot rate for Jan14
spotratejan14=c()
for(i in c(1:11))
  spotratejan14=append(spotratejan14,spotratebond[[i]][9])

###spot rate for Jan15
spotratejan15=c()
for(i in c(1:11))
  spotratejan15=append(spotratejan15,spotratebond[[i]][10])

###plot the spot curve
plot(ttmJan02, spotratejan02, type = "b", frame = FALSE, pch = 19, col="red", xlab = "years", ylab="spot rates")
lines(ttmJan03,spotratejan03, pch = 18, col = "blue", type = "b", lty = 1)
lines(ttmJan06, spotratejan06, pch = 18, col = "yellow", type = "b", lty = 1)
lines(ttmJan07, spotratejan07, pch = 18, col = "green", type = "b", lty = 1)
lines(ttmJan08, spotratejan08, pch = 18, col = "brown", type = "b", lty = 1)
lines(ttmJan09, spotratejan09, pch = 18, col = "purple", type = "b", lty = 1)
lines(ttmJan10, spotratejan10, pch = 18, col = "black", type = "b", lty = 1)
lines(ttmJan13, spotratejan13, pch = 18, col = "grey", type = "b", lty = 1)
lines(ttmJan14, spotratejan14, pch = 18, col = "pink", type = "b", lty = 1)
lines(ttmJan15, spotratejan15, pch = 18, col = "orange", type = "b", lty = 1)

legend("topright", legend=c("2020-01-02", "2020-01-03","2020-01-06","2020-01-07","2020-01-08","2020-01-09","2020-01-10","2020-01-13","2020-01-14","2020-01-15"),
       col=c("red", "blue","yellow","green","brown","purple","black","grey","pink","orange"), lty = 1, cex=0.5)


```




find the forward rate 
```{r setup, include=FALSE}
### find the forward rate
## use linear intepolation to calculate the spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff=c(yearFraction('2020-09-01','2021-03-01'),yearFraction('2021-09-01','2022-03-01'),yearFraction('2022-09-01','2023-03-01'),yearFraction('2023-09-01','2024-03-01'),yearFraction('2024-09-01','2025-03-01'))




### for Jan02, spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff02=c(yearFraction('2020-09-01','2021-01-02'),yearFraction('2021-09-01','2022-01-02'),yearFraction('2022-09-01','2023-01-02'),yearFraction('2023-09-01','2024-01-02'),yearFraction('2024-09-01','2025-01-02'))
spotrateyearlyJan02=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan02= append(spotrateyearlyJan02,spotratejan02[i]+(spotratejan02[i+1]-spotratejan02[i])/Timediff[i/2]*Timediff02[i/2])



### for Jan03,spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff03=c(yearFraction('2020-09-01','2021-01-03'),yearFraction('2021-09-01','2022-01-03'),yearFraction('2022-09-01','2023-01-03'),yearFraction('2023-09-01','2024-01-03'),yearFraction('2024-09-01','2025-01-03'))
spotrateyearlyJan03=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan03= append(spotrateyearlyJan03,spotratejan03[i]+(spotratejan03[i+1]-spotratejan03[i])/Timediff[i/2]*Timediff03[i/2])  



### for Jan06,spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff06=c(yearFraction('2020-09-01','2021-01-06'),yearFraction('2021-09-01','2022-01-06'),yearFraction('2022-09-01','2023-01-06'),yearFraction('2023-09-01','2024-01-06'),yearFraction('2024-09-01','2025-01-06'))
spotrateyearlyJan06=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan06= append(spotrateyearlyJan06,spotratejan06[i]+(spotratejan06[i+1]-spotratejan06[i])/Timediff[i/2]*Timediff06[i/2]) 



### for Jan07,spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff07=c(yearFraction('2020-09-01','2021-01-07'),yearFraction('2021-09-01','2022-01-07'),yearFraction('2022-09-01','2023-01-07'),yearFraction('2023-09-01','2024-01-07'),yearFraction('2024-09-01','2025-01-07'))
spotrateyearlyJan07=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan07= append(spotrateyearlyJan07,spotratejan07[i]+(spotratejan07[i+1]-spotratejan07[i])/Timediff[i/2]*Timediff07[i/2]) 



### for Jan08,spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff08=c(yearFraction('2020-09-01','2021-01-08'),yearFraction('2021-09-01','2022-01-08'),yearFraction('2022-09-01','2023-01-08'),yearFraction('2023-09-01','2024-01-08'),yearFraction('2024-09-01','2025-01-08'))
spotrateyearlyJan08=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan08= append(spotrateyearlyJan08,spotratejan08[i]+(spotratejan08[i+1]-spotratejan08[i])/Timediff[i/2]*Timediff08[i/2]) 



### for Jan09,spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff09=c(yearFraction('2020-09-01','2021-01-09'),yearFraction('2021-09-01','2022-01-09'),yearFraction('2022-09-01','2023-01-09'),yearFraction('2023-09-01','2024-01-09'),yearFraction('2024-09-01','2025-01-09'))
spotrateyearlyJan09=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan09= append(spotrateyearlyJan09,spotratejan09[i]+(spotratejan09[i+1]-spotratejan09[i])/Timediff[i/2]*Timediff09[i/2]) 



### for Jan10,spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff10=c(yearFraction('2020-09-01','2021-01-10'),yearFraction('2021-09-01','2022-01-10'),yearFraction('2022-09-01','2023-01-10'),yearFraction('2023-09-01','2024-01-10'),yearFraction('2024-09-01','2025-01-10'))
spotrateyearlyJan10=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan10= append(spotrateyearlyJan10,spotratejan10[i]+(spotratejan10[i+1]-spotratejan10[i])/Timediff[i/2]*Timediff10[i/2]) 



### for Jan13,spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff13=c(yearFraction('2020-09-01','2021-01-13'),yearFraction('2021-09-01','2022-01-13'),yearFraction('2022-09-01','2023-01-13'),yearFraction('2023-09-01','2024-01-13'),yearFraction('2024-09-01','2025-01-13'))
spotrateyearlyJan13=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan13= append(spotrateyearlyJan13,spotratejan13[i]+(spotratejan13[i+1]-spotratejan13[i])/Timediff[i/2]*Timediff13[i/2]) 



### for Jan14,spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff14=c(yearFraction('2020-09-01','2021-01-14'),yearFraction('2021-09-01','2022-01-14'),yearFraction('2022-09-01','2023-01-14'),yearFraction('2023-09-01','2024-01-14'),yearFraction('2024-09-01','2025-01-14'))
spotrateyearlyJan14=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan14= append(spotrateyearlyJan14,spotratejan14[i]+(spotratejan14[i+1]-spotratejan14[i])/Timediff[i/2]*Timediff14[i/2]) 



### for Jan15,spot rate in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff15=c(yearFraction('2020-09-01','2021-01-15'),yearFraction('2021-09-01','2022-01-15'),yearFraction('2022-09-01','2023-01-15'),yearFraction('2023-09-01','2024-01-15'),yearFraction('2024-09-01','2025-01-15'))
spotrateyearlyJan15=c()
for (i in c(2,4,6,8,10))
  spotrateyearlyJan15= append(spotrateyearlyJan15,spotratejan15[i]+(spotratejan15[i+1]-spotratejan15[i])/Timediff[i/2]*Timediff15[i/2]) 


###find the forward rate based on the spot rate in '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

### calculate forward rate for Jan02
forwardrateJan02=c()
for(i in c(2:5))
  forwardrateJan02=append(forwardrateJan02,((((1+spotrateyearlyJan02[i])^(i)/(1+spotrateyearlyJan02[1]))^(1/(i-1)))-1))

### calculate forward rate for Jan03
forwardrateJan03=c()
for(i in c(2:5))
  forwardrateJan03=append(forwardrateJan03,((((1+spotrateyearlyJan03[i])^(i)/(1+spotrateyearlyJan03[1]))^(1/(i-1)))-1))

### calculate forward rate for Jan06
forwardrateJan06=c()
for(i in c(2:5))
  forwardrateJan06=append(forwardrateJan06,((((1+spotrateyearlyJan06[i])^(i)/(1+spotrateyearlyJan06[1]))^(1/(i-1)))-1))

### calculate forward rate for Jan07
forwardrateJan07=c()
for(i in c(2:5))
  forwardrateJan07=append(forwardrateJan07,((((1+spotrateyearlyJan07[i])^(i)/(1+spotrateyearlyJan07[1]))^(1/(i-1)))-1))

### calculate forward rate for Jan08
forwardrateJan08=c()
for(i in c(2:5))
  forwardrateJan08=append(forwardrateJan08,((((1+spotrateyearlyJan08[i])^(i)/(1+spotrateyearlyJan08[1]))^(1/(i-1)))-1))

### calculate forward rate for Jan09
forwardrateJan09=c()
for(i in c(2:5))
  forwardrateJan09=append(forwardrateJan09,((((1+spotrateyearlyJan09[i])^(i)/(1+spotrateyearlyJan09[1]))^(1/(i-1)))-1))

### calculate forward rate for Jan10
forwardrateJan10=c()
for(i in c(2:5))
  forwardrateJan10=append(forwardrateJan10,((((1+spotrateyearlyJan10[i])^(i)/(1+spotrateyearlyJan10[1]))^(1/(i-1)))-1))

### calculate forward rate for Jan13
forwardrateJan13=c()
for(i in c(2:5))
  forwardrateJan13=append(forwardrateJan13,((((1+spotrateyearlyJan13[i])^(i)/(1+spotrateyearlyJan13[1]))^(1/(i-1)))-1))

### calculate forward rate for Jan14
forwardrateJan14=c()
for(i in c(2:5))
  forwardrateJan14=append(forwardrateJan14,((((1+spotrateyearlyJan14[i])^(i)/(1+spotrateyearlyJan14[1]))^(1/(i-1)))-1))

### calculate forward rate for Jan15
forwardrateJan15=c()
for(i in c(2:5))
  forwardrateJan15=append(forwardrateJan15,((((1+spotrateyearlyJan15[i])^(i)/(1+spotrateyearlyJan15[1]))^(1/(i-1)))-1))

### plot the forward curves
time=c(1,2,3,4)
plot(time, forwardrateJan02, type = "b", frame = FALSE, pch = 19, col="red", xlab = "years", ylab="forward rates",ylim = c(0.015,0.017))
lines(time,forwardrateJan03, pch = 18, col = "blue", type = "b", lty = 1)
lines(time, forwardrateJan06, pch = 18, col = "yellow", type = "b", lty = 1)
lines(time, forwardrateJan07, pch = 18, col = "green", type = "b", lty = 1)
lines(time, forwardrateJan08, pch = 18, col = "brown", type = "b", lty = 1)
lines(time, forwardrateJan09, pch = 18, col = "purple", type = "b", lty = 1)
lines(time, forwardrateJan10, pch = 18, col = "black", type = "b", lty = 1)
lines(time, forwardrateJan13, pch = 18, col = "grey", type = "b", lty = 1)
lines(time, forwardrateJan14, pch = 18, col = "pink", type = "b", lty = 1)
lines(time, forwardrateJan15, pch = 18, col = "orange", type = "b", lty = 1)

legend("topleft", legend=c("2020-01-02", "2020-01-03","2020-01-06","2020-01-07","2020-01-08","2020-01-09","2020-01-10","2020-01-13","2020-01-14","2020-01-15"),
       col=c("red", "blue","yellow","green","brown","purple","black","grey","pink","orange"), lty = 1, cex=0.4)
```




find Covarince Matrix
```{r setup, include=FALSE}


### for Jan02, yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.
Timediff02=c(yearFraction('2020-09-01','2021-01-02'),yearFraction('2021-09-01','2022-01-02'),yearFraction('2022-09-01','2023-01-02'),yearFraction('2023-09-01','2024-01-02'),yearFraction('2024-09-01','2025-01-02'))
yieldyearlyJan02=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan02= append(yieldyearlyJan02,Ytm02[i]+(Ytm02[i+1]-Ytm02[i])/Timediff[i/2]*Timediff02[i/2])

### for Jan03,yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff03=c(yearFraction('2020-09-01','2021-01-03'),yearFraction('2021-09-01','2022-01-03'),yearFraction('2022-09-01','2023-01-03'),yearFraction('2023-09-01','2024-01-03'),yearFraction('2024-09-01','2025-01-03'))
yieldyearlyJan03=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan03= append(yieldyearlyJan03,Ytm03[i]+(Ytm03[i+1]-Ytm03[i])/Timediff[i/2]*Timediff03[i/2])  



### for Jan06,yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff06=c(yearFraction('2020-09-01','2021-01-06'),yearFraction('2021-09-01','2022-01-06'),yearFraction('2022-09-01','2023-01-06'),yearFraction('2023-09-01','2024-01-06'),yearFraction('2024-09-01','2025-01-06'))
yieldyearlyJan06=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan06= append(yieldyearlyJan06,Ytm06[i]+(Ytm06[i+1]-Ytm06[i])/Timediff[i/2]*Timediff06[i/2]) 



### for Jan07,yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff07=c(yearFraction('2020-09-01','2021-01-07'),yearFraction('2021-09-01','2022-01-07'),yearFraction('2022-09-01','2023-01-07'),yearFraction('2023-09-01','2024-01-07'),yearFraction('2024-09-01','2025-01-07'))
yieldyearlyJan07=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan07= append(yieldyearlyJan07,Ytm07[i]+(Ytm07[i+1]-Ytm07[i])/Timediff[i/2]*Timediff07[i/2]) 



### for Jan08,yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff08=c(yearFraction('2020-09-01','2021-01-08'),yearFraction('2021-09-01','2022-01-08'),yearFraction('2022-09-01','2023-01-08'),yearFraction('2023-09-01','2024-01-08'),yearFraction('2024-09-01','2025-01-08'))
yieldyearlyJan08=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan08= append(yieldyearlyJan08,Ytm08[i]+(Ytm08[i+1]-Ytm08[i])/Timediff[i/2]*Timediff08[i/2]) 



### for Jan09,yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff09=c(yearFraction('2020-09-01','2021-01-09'),yearFraction('2021-09-01','2022-01-09'),yearFraction('2022-09-01','2023-01-09'),yearFraction('2023-09-01','2024-01-09'),yearFraction('2024-09-01','2025-01-09'))
yieldyearlyJan09=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan09= append(yieldyearlyJan09,Ytm09[i]+(Ytm09[i+1]-Ytm09[i])/Timediff[i/2]*Timediff09[i/2]) 



### for Jan10,yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff10=c(yearFraction('2020-09-01','2021-01-10'),yearFraction('2021-09-01','2022-01-10'),yearFraction('2022-09-01','2023-01-10'),yearFraction('2023-09-01','2024-01-10'),yearFraction('2024-09-01','2025-01-10'))
yieldyearlyJan10=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan10= append(yieldyearlyJan10,Ytm10[i]+(Ytm10[i+1]-Ytm10[i])/Timediff[i/2]*Timediff10[i/2]) 



### for Jan13,yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff13=c(yearFraction('2020-09-01','2021-01-13'),yearFraction('2021-09-01','2022-01-13'),yearFraction('2022-09-01','2023-01-13'),yearFraction('2023-09-01','2024-01-13'),yearFraction('2024-09-01','2025-01-13'))
yieldyearlyJan13=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan13= append(yieldyearlyJan13,Ytm13[i]+(Ytm13[i+1]-Ytm13[i])/Timediff[i/2]*Timediff13[i/2]) 



### for Jan14,yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff14=c(yearFraction('2020-09-01','2021-01-14'),yearFraction('2021-09-01','2022-01-14'),yearFraction('2022-09-01','2023-01-14'),yearFraction('2023-09-01','2024-01-14'),yearFraction('2024-09-01','2025-01-14'))
yieldyearlyJan14=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan14= append(yieldyearlyJan14,Ytm14[i]+(Ytm14[i+1]-Ytm14[i])/Timediff[i/2]*Timediff14[i/2]) 



### for Jan15,yieldcurve in the '2021-01-xx','2022-01-xx','2023-01-xx','2024-01-xx','2025-01-xx'.

Timediff15=c(yearFraction('2020-09-01','2021-01-15'),yearFraction('2021-09-01','2022-01-15'),yearFraction('2022-09-01','2023-01-15'),yearFraction('2023-09-01','2024-01-15'),yearFraction('2024-09-01','2025-01-15'))
yieldyearlyJan15=c()
for (i in c(2,4,6,8,10))
  yieldyearlyJan15= append(yieldyearlyJan15,Ytm15[i]+(Ytm15[i+1]-Ytm15[i])/Timediff[i/2]*Timediff15[i/2]) 

yieldyear=c(yieldyearlyJan02,yieldyearlyJan03,yieldyearlyJan06,yieldyearlyJan07,yieldyearlyJan08,yieldyearlyJan09,yieldyearlyJan10,yieldyearlyJan13,yieldyearlyJan14,yieldyearlyJan15)


yieldyear = matrix(yieldyear,  nrow = 10, ncol = 5, byrow = T)
yieldyear=t(yieldyear)

Xyield = c()
for (i in c(1:5)){
  for (j in c(1:9)){
   Xyield = append(Xyield, log(yieldyear[i,j+1]/Yield[i,j]))
  }
}
Xyield = matrix(Xyield, nrow = 5, ncol = 9, byrow = T)
### Covariance Matrix of Yield log return and eigenvectors & eigenvalues
cov_yield = cov(t(Xyield))
cov_yield
eigenyield=eigen(cov_yield)
eigenyield




##### covariance of forward rate
forwardrate=c(forwardrateJan02,forwardrateJan03,forwardrateJan06,forwardrateJan07,forwardrateJan08,forwardrateJan09,forwardrateJan10,forwardrateJan13,forwardrateJan14,forwardrateJan15)
forwardrate=matrix(forwardrate, nrow=10, ncol=4, byrow = T)

Xforward = c()
for (i in c(1:4)){
  for (j in c(1:9)){
    Xforward = append(Xforward, log(forwardrate[j+1,i]/forwardrate[j,i]))
  }
}
Xforward = matrix(Xforward, nrow = 4, ncol = 9, byrow = T)
### Covariance Matrix of Forward log return and eigenvectors & eigenvalues
cov_forward = cov(t(Xforward))
cov_forward
eigenforward=eigen(cov_forward)
eigenforward